---
title: "insuranceDemo"
author: "Daniel"
date: "March 16, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r echo=FALSE}
rstudioapi::getActiveDocumentContext()
this.dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(this.dir)

#Suppressing Warning so they do not print when knitting
#Do not not run when testing
oldw <- getOption("warn")

#UNCOMMENT ONLY WHEN KNITING
options(warn = -1)

#need to include at the bottom of script
#options(warn = oldw)

```

```{r libraries, echo=FALSE}
library(Matrix)
library(glmnet)
library(mlbench)
library(caret)
library(ggplot2)
library(GGally)
library(nnet)
library(NeuralNetTools)
library(dummies)

```

## Loading Data

```{r data, echo=FALSE}

data <- read.csv("./customer_claims_demo.csv")

head(data)

```

## Data Exploration

Summaries of Data

###Total Claim Amount

``` {r claim histogram, echo = FALSE}

library(ggplot2)

qplot(data$Total.Claim.Amount, geom = "histogram")


```

###Income

``` {r Income historgram, echo=FALSE}


qplot(data$Income, geom = "histogram")


```

Many Zero Valuse, presumably uemployed, use colors to confirm

``` {r visuals, echo=FALSE}

qplot(data$Income, geom = "histogram"
      , fill = data$EmploymentStatus)


```


### Scatter Matrix

Comapring pairwise distribution of all continuous values

Monthly Premium, CLV, and Income show some correlation with Total Claim amount, however it looks like Employment Status remains a confounding factor

```{r scatter matrix color, echo = TRUE, message = FALSE, warning= FALSE}

library(GGally)

numerics <- sapply(data, is.numeric) #Get numeric columns only

dataCorrs <- data[,numerics] #Filter data to only numeric for correlations

#Create new value 'EmploymentStatus' with all employed as 'Employed' otherwise unemployed
dataCorrs$EmploymentStatus <- ifelse(data$EmploymentStatus == 'Employed', 'Employed', 'Unemployed')

ggpairs(dataCorrs
                , mapping = ggplot2::aes(color= EmploymentStatus)
                , lower=list(combo=wrap("facethist",binwidth=10))
                , upper=list(continuous=wrap("cor", size=2))
        )





```




## Model Evaluation

Using Caret to evaluate Ridge Lasso and Linear regression options

```{r caret, message= = FALSE}

dataFactors <- data.frame(Education = data$Education
                          ,EmploymentStatus = data$EmploymentStatus
                          ,Location.Code = data$Location.Code
                          ,Policy.Type = data$Policy.Type
                          ,Vehicle.Class = data$Vehicle.Class
                          ,Vehicle.Size = data$Vehicle.Size)


#adding to check distribution
dataFactors$Total.Claim.Amount <- data$Total.Claim.Amount


#sample data set

dataNumeric$EmploymentStatus <- NULL
dataNumeric$Total.Claim.Amount <- NULL

dataset <- data.frame(dataFactors, dataNumeric )


cv.train <- createDataPartition(dataset$Total.Claim.Amount, p = 0.5, list = FALSE)

dataset.train <- dataset[cv.train, ]
dataset.test <- dataset[-cv.train, ]


#Build Formula

names <- names(dataset)
y <- "Total.Claim.Amount"
x <- names[!names %in% y]
form <- as.formula(paste(y, " ~ ",paste(x,collapse="+")))

glm_results <- NULL

glm_results <- caret::train(form,
                               data = dataset.train,
                               method = "glmnet",
                               metric = "Rsquared",
                               tuneGrid = expand.grid(alpha=c(0,0.01,0.05,0.99,1),lambda = c(0,0.01)), 
                               trControl = trainControl(method="cv"),
                               preProcess = c("center", "scale"),
                               na.action=na.omit
)


```

```{r plotting results 1}
plot(glm_results)
```

```{r plotting results 2}

#Get coef names
vnat=coef(glm_results$finalModel)
vnat=vnat[-1,ncol(vnat)] # remove the intercept, and get the coefficients at the end of the path
vn = names(vnat)

plot(glm_results$finalModel, xvar="lambda", label = TRUE)
axis(4, at=vnat,line=-.5,label=vn,las=1,tick=FALSE, cex.axis=0.5) 
```

```{r plotting results 3}
plot(glm_results$finalModel, xvar="dev", label= TRUE)
axis(4, at=vnat,line=-.5,label=vn,las=1,tick=FALSE, cex.axis=0.5) 
```

```{r plotting results 4}
glm_test.lmin <- predict(glm_results, dataset.test, s="lambda.min", type = "raw")

head(glm_test.lmin)
smoothScatter(glm_test.lmin, dataset.test$Total.Claim.Amount)
```



Using base glmnet package now that we've seen Alpha = 1 is optimal


```{r glmnet cv}

y.cv <- data[,names(data) %in% y]

x.cv <- as.matrix(dataNumeric[,!names(dataNumeric) %in% y])

glm_cv <- cv.glmnet(x = x.cv,y = y.cv)

plot(glm_cv)

coef(glm_cv, s="lambda.min")


coef(glm_cv, s="lambda.1se")

```

```{r glmnet cv 2}

data.oh <- data

data.oh$Customer <- NULL

names.all <- names(data.oh)

names.factors <- names.all[!names.all %in% names(dataNumeric)]

names.factors <- names.factors[names.factors!=y]

names.factors <- names.factors[names.factors!="Effective.To.Date"]

data.ohc <- dummy.data.frame(data.oh, names=names.factors, sep=".")

y.cvoh <- data.ohc[,names(data.ohc) %in% y]

x.cvoh <- data.ohc[,!names(data.ohc) %in% y]

x.cvoh <- as.matrix((sapply(x.cv,as.numeric)))

glm_cvoh <- cv.glmnet(x = x.cvoh,y = y.cvoh)

plot(glm_cvoh)

coef(glm_cvoh, s="lambda.min")

```


#Tree diagram

A regression or decision tree shows a heirarchy of factors influencing some outcome.

Numeric values are split (high/low) at the point creating the greatest impact on outcome value.

Multiple factors are evaluated as one vs any other a.k.a. 'one-hot' or 'dummy values'

```{r}

library(rpart)
library(rpart.plot)

head(data)

tree = rpart(Total.Claim.Amount ~ Vehicle.Size + Policy.Type + Months.Since.Last.Claim + Marital.Status + Income + EmploymentStatus
             , method = "anova"
             , data = data)

rpart.plot(tree)


```

Trees give us a nice view of what influences our outcome

We can get a high level view of our data quickly...

```{r tree plot full}

library(rpart)
library(rpart.plot)

treeFormula = as.formula(
  "Total.Claim.Amount ~  Coverage + Education + EmploymentStatus 
  + Gender + Income + Location.Code + Marital.Status + Vehicle.Size"
  )

treeAll = rpart(treeFormula
             , method = "anova"
             , data = data)

rpart.plot(treeAll)


```

Then do some more exploratory analysis to get a better picture

```{r tree validation}

library(ggplot2)

ggplot(data, aes(x=Total.Claim.Amount, fill=Location.Code)) 
  + geom_density(alpha = .5)

ggplot(data, aes(x=Total.Claim.Amount, fill=Coverage)) 
  + geom_density(alpha = .5)


```

```{r normalizing}

data$Claim.To.Premium = data$Total.Claim.Amount / data$Monthly.Premium.Auto

ggplot(data, aes(x=Claim.To.Premium, fill=Location.Code)) 
  + geom_density(alpha = .5)

ggplot(data, aes(x=Claim.To.Premium, fill=Coverage)) 
  + geom_density(alpha = .5)

ggplot(data, aes(x=Claim.To.Premium, fill=Marital.Status)) + geom_density(alpha = .5)

ggplot(data, aes(x=Claim.To.Premium, fill=Income.Over.5137)) + geom_density(alpha = .5)
       

```

```{r tree normalized}

treeNorm = rpart( 
             , method = "anova"
             , data = data)

rpart.plot(treeNorm)

```


```{r Marital Status}


ggplot(data, aes(x=Claim.To.Premium, fill=Marital.Status)) + geom_density(alpha = .5)

data$Income.Over.5137 = ifelse(data$Income > 5137, TRUE, FALSE)

ggplot(data, aes(x=Claim.To.Premium, fill=Income.Over.5137)) + geom_density(alpha = .5)

# Getting fancy

library(dplyr)

data = data %>% 
  mutate(
    Location.Number = case_when(
    Location.Code == "Rural" ~ 0,
    Location.Code == "Suburban" ~ 1,
    Location.Code == "Urban" ~ 2,
    TRUE ~ 4
  )
)

head(data)

library(plotly)

pData = data[data$Income > 0,]

p = plot_ly(pData, x = ~Income, y = ~Claim.To.Premium, z = ~Monthly.Premium.Auto, color = ~Location.Code) %>% add_markers()

p

library(reshape2)

mData = (data[, names(data) %in% c("Location.Number", "Total.Claim.Amount", "Income")])

mData = melt(mData, id.vars = c("Location.Number", "Income"), measure.vars ="Total.Claim.Amount")

mData$variable = NULL


head(mData)

summary(mData)


ggplot(mData, aes(z = Location.Number, y = Income, x = value)) + stat_contour()



```
#Build NNET Formula

```{r nn, message=FALSE}

names <- names(dataNumeric)
y_nn <- "Total.Claim.Amount"
x_nn <- names[!names %in% y]
form_nn <- as.formula(paste(y_nn, " ~ ",paste(x_nn,collapse="+")))

nn1_grid <- expand.grid(decay = c(0.01, .1), size = c(3, 5))

nn1_preProc <- c("center","scale")


nn1 <- NULL

nn1 <- train(form_nn, 
             data=dataset.train, 
             method="nnet", 
             linout=TRUE, 
             tuneGrid = nn1_grid, 
             preProcess = nn1_preProc,
             maxit = 100,
             metric = "Rsquared",
             maximize = TRUE)


plot(nn1)

plotnet(nn1$finalModel)

lekprofile(nn1, xsel = c('Customer.Lifetime.Value', 'Income'), group_vals = seq(0,1,by=0.05))


```

#Additional Testing with Earth

Just for visualizations

```{r caret earth, message= = FALSE}

#sample data set

library(earth)
library(caret)

numerics <- sapply(data, is.numeric)

dataNumeric <- data[,numerics]

dataNumeric$EmploymentStatus <- ifelse(data$EmploymentStatus == 'Employed', 'Employed', 'Unemployed')


dataFactors <- data.frame(Education = data$Education
                          ,EmploymentStatus = data$EmploymentStatus
                          ,Location.Code = data$Location.Code
                          ,Policy.Type = data$Policy.Type
                          ,Vehicle.Class = data$Vehicle.Class
                          ,Vehicle.Size = data$Vehicle.Size)


dataNumeric$EmploymentStatus <- NULL
dataNumeric$Total.Claim.Amount <- NULL

dataFactors$Total.Claim.Amount <- data$Total.Claim.Amount

dataset <- data.frame(dataFactors, dataNumeric )


cv.train <- createDataPartition(dataset$Total.Claim.Amount, p = 0.5, list = FALSE)

dataset.train <- dataset[cv.train, ]
dataset.test <- dataset[-cv.train, ]


#Build Formula

names <- names(dataset)
y <- "Total.Claim.Amount"
x <- names[!names %in% y]
form <- as.formula(paste(y, " ~ ",paste(x,collapse="+")))

form2 <- as.formula("Claim.To.Premium ~ Customer.Lifetime.Value + Coverage + Education + EmploymentStatus + Gender + Income + Location.Code + Marital.Status + Vehicle.Size")

earth_results <- NULL

earth_results <- earth(form2,
                       data = data,
                       degree = 2,
                       nprune = 5
                      
)


```

#plotting earth

```{r caret earth visuals, message = FALSE}

plotmo(earth_results)

#head(dataset)


```
